---
title: "DIStudio Quick Start"
author: "Ryan Barnard"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DIStudio Quick Start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

First, we load the DIStudio package and load the default data model:
```{r}
library(DIStudio)

model <- loadDefaultModel()
```

Next, suppose we're interested in Alzheimer's Disease (AD) drugs. We can lookup related nodes:
```{r}
listNodes(model, name="alzheimer")
```

The list can also be filtered to include only diseases:
```{r}
listNodes(model, name="alzheimer", kind=DISEASE_KIND)
```

From here, we have the code for Alzheimer's Disease: `C904`. To find drugs related to AD, use the `listRelatedDrugs`
function:
```{r}
drugs <- listRelatedDrugs(model, "C904")
nrow(drugs)
head(drugs)
```

To lookup annotations related to one of these drugs, use the `listDrugAnnotations` function:
```{r}
annotations <- listDrugAnnotations(model, "C14004")
nrow(annotations)
head(annotations)
```

The list can be restricted to specific annotation kinds:
```{r}
listDrugAnnotations(model, "C14004", kinds=MECHANISM_OF_ACTION_KIND)
```


Drug similarity can be calculated with the `Similarity` function:
```{r}
Similarity(model, c("C14004", "C13950"), "C14954")
```

Finally, the model can be visualized with the `viewModel` function. The entire default model is too large
to be meaningfully graphed, but the `getDataSubmodel` function can extract a portion of
the model containing only a subset of nodes:
```{r}
codes <- c(drugs$code, annotations$code)
submodel <- getDataSubmodel(model, codes)

viewModel(submodel)
```
