---
title: "rxnorm-example"
author: "Shyh-Huei Chen"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RxNorm Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

In this example, we will demonstrate some essential functionalities related to
graph-based similarities. We used simulated electronic medical records of the
drug use of patients with glaucoma compared with patients with AIDS. 

First, we load the DIStudio package and load the default data model:
```{r}
library(DIStudio)

model <- loadDefaultModel()
```

Then we show different means to explore drugs of interest with keywords. One frequently
used strategy is to query the NDF-RT concepts with molecular features. It is known that
carbonic anhydrase inhibitors are widely used to treat glaucoma. Searching the NDF-RT
graph model with keyword “carbonic” will list related concepts: 

```{r}
listNodes(model, name='carbonic')
```

Concepts about Mechanism of Action (MoA, kind C12), chemical or ingredient (kind C10),
etc., are listed. We further explore drugs annotated by the MoA concept “Carbonic
Anhydrase Inhibitors” (with the concept code “C648” and concept kind as “C12”) and
visualize the ontology network of these drugs: 

```{r fig.height=6}
carbonicNodes <- listRelatedNodes(model, 'C648', direction='both')
viewModel(getDataSubmodel(model, carbonicNodes$code))
```

Another strategy is to explore by disease (“HIV”): 
```{r}
listNodes(model, name='HIV')
```

And further explore this list for concepts of interest. For example, we extract drugs
annotated by the MoA concept “HIV Protease Inhibitors” (“C670”) and visualize the ontology
of these drugs:
```{r fig.height=6}
hivNodes <- listRelatedNodes(model, 'C670', direction='both')
viewModel(getDataSubmodel(model, hivNodes$code))
```

In the next section, we will demonstrate functionality related to similarity measures
using two simulated cohorts, the glaucoma cohort (cohort 648) of 500 patients with drug
records enriched with the MoA concept “Carbonic Anhydrase Inhibitors” (C648), and the AIDS
cohort (cohort 670) of 500 patients with drug records enriched with the MoA concept “HIV
Protease Inhibitors” (“C670”). The drug records of each patient are represented by a list
of RxNorm codes. 

Load the RxNorm to NDF-RT mapping and some sample data:
```{r}
rxnorm <- loadRxNormMapping()
data('rxNormCodes')
names(rxNormCodes)
```

The list `Cohort.Glaucoma.original` of 500 elements represent 500 patients in the glaucoma cohort. The drug records of the first patient in this cohort, coded by RxNorm code, are: 

```{r}
rxNormCodes$Cohort.Glaucoma.original[[1]]
```

Similarly, the drug records of the first patient in the AIDS cohort, coded by RxNorm code,
are:

```{r}
rxNormCodes$Cohort.AIDS.original[[1]]
```

To analyze such drug records, we first need to map RxNorm codes to NDF-RT codes. We
provided a manually curated mapping table for this purpose. 

```{r}
# A simple helper function to simplify conversion of RxNorm CUIs to NDF-RT concept codes.
cui2concept <- function(x, mapping) {
   y <- mapping$concept_code[match(x, mapping$RX_CUI)]
   y <- y[!is.na(y)]

   return(y)
}

Cohort.Glaucoma.concept <- lapply(rxNormCodes$Cohort.Glaucoma.original, cui2concept, rxnorm)
Cohort.AIDS.concept  <- lapply(rxNormCodes$Cohort.AIDS.original, cui2concept, rxnorm)
```

Then we demonstrate the following similarity measures and visualization: 
```{r}
Cohort.Glaucoma.Patient1 <-  Cohort.Glaucoma.concept[[1]]
Cohort.Glaucoma.Patient2 <-  Cohort.Glaucoma.concept[[2]] 

Cohort.AIDS.Patient1 <-  Cohort.AIDS.concept[[1]]  
Cohort.AIDS.Patient2 <-  Cohort.AIDS.concept[[2]] 
```

1) Comparing the similarity of two drugs: these two drugs, “C9716” and “C35700”, are
different in medical records. However, since these two drugs share identical MoA
annotations, they are highly similar once projected onto the MoA knowledge domain and have
similarity of 1. 
```{r}
Cohort.Glaucoma.Patient1[1]
Cohort.Glaucoma.Patient2[2]
Similarity(model, Cohort.Glaucoma.Patient1[1], Cohort.Glaucoma.Patient2[2], View=T)
```


2) Comparing the similarity of two drugs: these two drugs, “C9716” and “C50.68.17991”, are
less similar once projected onto the MoA knowledge domain and have similarity of 0.5. This
similarity can be comprehended by the overlapping of their MoA annotation on the graph, as
demonstrated in the plot. 

```{r}
Cohort.Glaucoma.Patient1[1]
Cohort.AIDS.Patient1[2]
Similarity(model, Cohort.Glaucoma.Patient1[1], Cohort.AIDS.Patient1[2], View=T)
```

3) Comparing the similarity of drug records between two patients (or between any two lists
of drug records). These two patients are from different cohorts and have different
clinical conditions (glaucoma vs. AIDS). Such difference can be captured by the lower
similarity between the drugs they used (0.1315789). The similarity can be comprehended by
the overlapping of their MoA annotation on the graph, as demonstrated in the plot. 

```{r}
Cohort.Glaucoma.Patient1
Cohort.AIDS.Patient1
Similarity(model, Cohort.Glaucoma.Patient1, Cohort.AIDS.Patient1, View=T)
```

Another way to compare a pair of drug lists is to calculate their frequencies and
compute the Jaccard similarity between the frequencies:
```{r}
submodel <- getKindsSubmodel(model, c(DRUG_KIND, MECHANISM_OF_ACTION_KIND))

Cohort.Glaucoma.Patient1.Freq <- Frequencies(submodel, Cohort.Glaucoma.Patient1)
Cohort.Glaucoma.Patient2.Freq <- Frequencies(submodel, Cohort.AIDS.Patient1)

Jaccard.F(Cohort.Glaucoma.Patient1.Freq, Cohort.Glaucoma.Patient2.Freq)
```

These drug lists can be also be visualized by creating subgraphs of the full model 
consisting of nodes from our simulated patients' drug lists:
```{r}
# Add the counts together to get the nodes in either Cohort.Glaucoma.Patient1.Freq or Cohort.Glaucoma.Patient2.Freq
countF1F2 <- Cohort.Glaucoma.Patient1.Freq$vertices$count + Cohort.Glaucoma.Patient2.Freq$vertices$count

F1.G <- igraph::induced_subgraph(model$graph, with(Cohort.Glaucoma.Patient1.Freq$vertices, code[count > 0]))
F2.G <- igraph::induced_subgraph(model$graph, with(Cohort.Glaucoma.Patient2.Freq$vertices, code[count > 0]))
F1F2.G <- igraph::induced_subgraph(model$graph, Cohort.Glaucoma.Patient1.Freq$vertices$code[countF1F2 > 0])
```

Once constructed, the subgraphs themselves can be visualized with the `plot` function:
```{r}
plot(F1.G)
plot(F2.G)
plot(F1F2.G)
```

The subgraph of both patient's drugs combined can be plotted with the different layout
functions available in the `igraph` package.
```{r}
rxNormPalette <- colorRampPalette(c("red", "yellow", "green"))(n = 101)

# Set the size and color of the vertices
igraph::V(F1F2.G)$size <- ceiling(log10(countF1F2[countF1F2>0]+1)*5)
igraph::V(F1F2.G)$color <- rxNormPalette[(round(Cohort.Glaucoma.Patient1.Freq$vertices$count[countF1F2>0]/countF1F2[countF1F2>0]*100)+1)]

# A tree plot makes it easy to see the relationships, but it's not elegant 
plot(F1F2.G, vertex.color=igraph::V(F1F2.G)$color, layout=igraph::layout_as_tree(F1F2.G, mode = "in"))
# Large graph layout looks nice, but it's hard to see the similarity/dissimilarity
plot(F1F2.G, vertex.color=igraph::V(F1F2.G)$color, layout=igraph::layout_with_lgl(F1F2.G))
```
